Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-06 ms.tex
added to definitions table

In the presence of strong heating, one-dimensional steady-state flow
is characterized by an inflow-outflow structure, with a critical
radius known as the``stagnation radius" $\rs$ where the radial
velocity passes through zero.  Mass loss from stars interior to the
stagnation radius is accreted, while that outside $\rs$ is unbound in
an outflow from the nucleus.  The stagnation radius, rather than the
Bondi radius, thus controls the SMBH accretion rate (although we will
show that $\rs$ can in many cases lie within a factor of few of the
nominal Bondi radius).  In the case of sufficiently weak heating,
however, no stagnation radius may exist at all, i.e. the entire ISM of
the galaxy is feeding the SMBH.  Another goal of this work is to
assess under what conditions a stagnation radius exists, as this
question is intimately tied to that of cooling instability.
% Furthermore, in some cases we find that no stagnation
% radius exists at all; this formally implies that the entire of ISM of
% the galaxy is feeding the SMBH, potentially resulting in much higher
% accretion rate than would be predicted by simple Bondi accretion.

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe some
analytic results, while in $\S\ref{sec:numerical}$ we describe our numerical results.  In $\S\ref{sec:discussion}$ we discuss the implications of our results, to topics including...   In $\S\ref{sec:conclusions}$ we summarize our conclusions.  Table \ref{table:definitions} provides the definitions of commonly used variables.  Appendix \ref{app:rs} provides useful analytic results, while Appendix \ref{app:windheat} describes our method for calculating stellar wind heating and mass input in detail.

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{lc}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & Black hole mass \\
$\tilde{v}_{w}$ & Total heating parameter, including minimum heating rate from stellar velocity dispersion \\
$v_{w}$ & Total heating parameter, excluding heating rate from stellar velocity dispersion \\
$r_{s}$ & Stagnation radius (where radial velocity goes to zero) \\
$r_{\rm soi}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\  
\hline
\label{table:definitions}  
\end{tabular}
\begin{tablenotes}
\item{$^{(a)}FILL IN$}
\end{tablenotes}
\end{minipage}
\end{threeparttable}

>>>>

In the presence of strong heating, one-dimensional steady-state flow
is characterized by an inflow-outflow structure, with a critical
radius known as the``stagnation radius" $\rs$ where the radial
velocity passes through zero.  Mass loss from stars interior to the
stagnation radius is accreted, while that outside $\rs$ is unbound in
an outflow from the nucleus.  The stagnation radius, rather than the
Bondi radius, thus controls the SMBH accretion rate (although we will
show that $\rs$ can in many cases lie within a factor of few of the
nominal Bondi radius).  In the case of sufficiently weak heating,
however, no stagnation radius may exist at all, i.e. in principle the entire ISM of
the galaxy is feeding the SMBH.  Another goal of this work is to
assess under what conditions a stagnation radius exists, as this
question is intimately tied to that of cooling instability.
% Furthermore, in some cases we find that no stagnation
% radius exists at all; this formally implies that the entire of ISM of
% the galaxy is feeding the SMBH, potentially resulting in much higher
% accretion rate than would be predicted by simple Bondi accretion.

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe some
analytic results, while in $\S\ref{sec:numerical}$ we describe our numerical results.  In $\S\ref{sec:discussion}$ we discuss the implications of our results, to topics including...   In $\S\ref{sec:conclusions}$ we summarize our conclusions.  Table \ref{table:definitions} provides the definitions of commonly used variables.  Appendix \ref{app:rs} provides useful analytic results, while Appendix \ref{app:windheat} describes our method for calculating stellar wind heating and mass input in greater detail.

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{lc}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & Black hole mass \\
$\tilde{v}_{w}$ & Total heating parameter, including minimum heating rate from stellar velocity dispersion \\
$v_{w}$ & Total heating parameter, excluding heating rate from stellar velocity dispersion \\
$r_{s}$ & Stagnation radius (where radial velocity goes to zero) \\
$r_{\rm soi}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\  
$\rho_{\star}(r)$ & 3D radial stellar density profile \\
$\rho(r)$ & gas density of CNM \\
$M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
$M_{\rm enc}(r)$ & total encolsed mass inside radius $r$ (SMBH + stars) \\
$q(r)$ & mass source term due to stellar winds (eq.~[\ref{eq:q}]) \\
$\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
$\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
$\Gamma$ & power-law slope of radial stellar density profile interior to the break radius \\
$\sigma(r)$ & stellar velocity dispersion \\
$\dot{M}$ & mass accretion rate \\
$\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]) \\
$\dot{M}_{\rm C}$ & equilibrium accretion rate set by Compton heating (eq.~[\ref{eq:MdotC}]) \\
$\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
$r_{\rm Ia}$ & radius exterior to which Ia SN are infrequent compared to the dynamical/inflow timescale (eq.~[\ref{eq:rIa}]) \\
\hline
\label{table:definitions}  
\end{tabular}
\begin{tablenotes}
\item{$^{(a)}FILL IN$}
\end{tablenotes}
\end{minipage}
\end{threeparttable}
